---
title: "Homework 1"
author: "Christine Catherman"
date: "Februrary 3rd, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

## Task 1

```{r mh_distance_function}
mh_distance <- function(x, y) {
  
#Create a list of x and y
  x_char <- unlist(strsplit(as.character(x), NULL))
  y_char <- unlist(strsplit(as.character(y), NULL))
  
  #If numeric then take the absolute value
  if (is.numeric(x) & is.numeric(y)) {
    x_pos <- abs(x)
    y_pos <- abs(y)
    x_char <- unlist(strsplit(as.character(x_pos), NULL))
    y_char <- unlist(strsplit(as.character(y_pos), NULL))
  }
  
 # Check Inputs
  #Both same types
  if (typeof(x) != typeof(y)) {
  warning("x and y are not both logical, numeric, or characters")
  return(-1)
  }
  
  # Same Length
  if (typeof(x) != "logical" & typeof(y) != "logical") {
  if (nchar(x) != nchar(y)) {
  warning("x and y do not have the same number of digits or letters")
  return(-1)
  }
  }
  
  #Decimals
  if ( is.numeric(x) & is.numeric(y)) {
    for(i in x_char) {
      if (i == ".") {
        warning("x contains decimal values")
    return (-1)
      }
    }
  
   for(l in y_char) {
      if (l == ".") {
        warning("y contains decimal values")
    return (-1)
      }
    }
  }
  
  #NA, NaN, Inf or -Inf
  if (is.infinite(x) | is.nan(x) | is.na(x)) {
  warning("x is NA, NaN, Inf or -Inf")
  return(-1)
    }
  
  if (is.infinite(y) | is.nan(y) | is.na(y)) {
  warning("y is NA, NaN, Inf or -Inf")
  return(-1)
   }

  
#Run Calculations
  #Character or Numeric
  if ((is.character(x) & is.character(y)) || (is.numeric(x) & is.numeric(y))) {
    count <- 0
    i <- 1
      while(i <= length(x_char)) {
      if (x_char[i] != y_char[i]) {
        count = count + 1
      }
        i <- i + 1
    }
    return (count)
  }

  #Logical
  if (is.logical(x) & is.logical(y)) {
    value <- abs(x - y)
      return (value)
  }
  
}
```


## Task 2

<i>
Be sure to remove chunk option `eval=FALSE` in the below code chunks
once you create `mh_distance()`. Do not include this italicized
text in your final submission.
</i>

**Initial test cases** (return a non-negative modified Hamming distance)

```{r}
mh_distance(x = "abc", y = "abc")
mh_distance(x = T, y = FALSE)
mh_distance(x = "523890", y = "752839")
mh_distance(x = 2341, y = 2350)
```

Added test cases that return a non-negative modified Hamming distance result.

```{r}
# additional test cases go here
mh_distance(x = FALSE, y = TRUE)
mh_distance(x = "cac", y = "abc")
mh_distance(x = "dog", y = "dog")
mh_distance(x = -5, y = 5) #Still not working!!! with negative numbers
```

**Initial test cases** (return values of `-1`)

```{r eval=FALSE}
mh_distance(x = 52, y = 113)
mh_distance(x = "swimming", y = "winning")
mh_distance(x = NA, y = TRUE)
mh_distance(x = 1.5, y = 2.5)
```

Added test cases that return a value of `-1`.

```{r}
# additional test cases go here
mh_distance(x = TRUE, y = 2.5)
mh_distance(x = 5, y = "cac")
mh_distance(x = "cac", y = 2)
mh_distance(x = "cac", y = "christine")
mh_distance(x = 123, y = 2)
mh_distance(x = Inf, y = 500)
mh_distance(x = FALSE, y = NA)
mh_distance(x = 671, y = 5.3)
```


## Task 3

Consider the pair of vectors `s` and `w` given below.

```{r}
s <- c(26, 50123, 456.12, 8, 0)
w <- c(22, 50000, 451.00, 88, 0)
```

```{r}
# loop goes here

```


## Task 4

Brief write-up goes here...
